% ============================================================
% EECE 693 – Neural Networks / Deep Learning – Progress Report
% ============================================================
\documentclass[11pt,a4paper]{article}

% ── packages ──────────────────────────────────────────────────
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{float}
\usepackage[numbers]{natbib}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{array}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=blue!60!black,
  urlcolor=blue!60!black
}

% ── title ─────────────────────────────────────────────────────
\title{%
  \textbf{EECE 693 (Neural Networks / Deep Learning) – Progress Report}\\[6pt]
  \Large Early Warning Prediction of Asthma Exacerbations\\
  Using Wearable and Digital-Health Data}

\author{%
  Bahaa Hamdan \qquad Tarek El Mourad \qquad Anthony Kanaan \\[4pt]
  Department of Electrical and Computer Engineering,\\
  American University of Beirut (AUB) \qquad Spring 2026\\
  Beirut, Lebanon}

\date{}

\begin{document}
\maketitle

% ============================================================
\begin{abstract}
% ============================================================
Asthma exacerbations are acute episodes of worsening respiratory symptoms
that impose substantial clinical and economic burden worldwide, and early
detection from continuous wearable signals could in principle enable timely
intervention.  Whether such detection is feasible at a small-cohort scale
is an open question that this work addresses negatively.  We present an
end-to-end pipeline for predicting asthma exacerbations from the AAMOS-00
study (22 enrolled, 18 contributing usable labelled windows, ${\sim}2{,}054$
patient-days), evaluated under a 5-seed protocol with patient-cluster
bootstrap 95\% confidence intervals, an automated three-probe leakage
audit, and Platt-calibrated outputs for the deep models.  The pipeline
spans data cleaning, weekly-questionnaire OR-of-flags labelling (30.3\%
positive rate, 31{,}124 windows), Tier-1 tabular baselines, four Tier-2
sequence models (GRU, LSTM, vanilla tanh-RNN, 3-layer 1D~CNN), and a
Tier-3 multimodal architecture\,$\times$\,modality sweep
(3~recurrent backbones $\times$ 4~cumulative co-modality subsets
$\times$ 5~seeds).  \textbf{The principal empirical finding is a
non-result.}  All four Tier-2 deep models sit at chance (multi-seed
ROC-AUC 0.46--0.51, every per-seed bootstrap interval crossing 0.5).  The
strongest multimodal configuration (LSTM with smartwatch + smart-inhaler
+ peak-flow) reaches ROC-AUC\,=\,0.634\,$\pm$\,0.062, but its per-seed
bootstrap lower bound dips below 0.5 on individual seeds and is therefore
not statistically distinguishable from chance.  Seven candidate
improvements (per-patient normalisation, stratified-group split,
LightGBM/XGBoost on the multimodal panel, hybrid Tier-3 fusion, and a
three-learner stacking ensemble) were tested under the same protocol;
\emph{none beat the baseline} (best contender
0.625\,$\pm$\,0.059).  A simple power calculation
(Section~\ref{sec:limitations}) confirms that detecting a true
AUC$\approx$0.63 against chance with~$18$ patients is below conventional
power; the minimum AUC reliably detectable at this cohort size is
${\approx}0.75$.  Our contribution is therefore methodological rather
than predictive: we provide an evaluation protocol for small-cohort
wearable studies that is robust to seed variance and label leakage, and
we document, transparently, that this dataset is too small to support
clinically meaningful exacerbation prediction.
\end{abstract}

\tableofcontents
\newpage

% ============================================================
\section{Introduction}\label{sec:intro}
% ============================================================

Asthma is a chronic inflammatory airway disease affecting an estimated
262~million people worldwide and contributing to approximately 455{,}000 deaths
annually~\cite{who2024}.  Exacerbations—acute or sub-acute episodes of
worsening symptoms and/or lung-function decline—represent the most
consequential events in asthma care.  They can lead to emergency visits,
hospitalization, and life-threatening deterioration if not recognized
early~\cite{bourdin2019}.  The total annual burden in the United States alone
has been estimated at~\$81.9~billion~\cite{nurmagambetov2018}.

Evidence shows that outcomes improve when worsening is identified early and
patients act on structured self-management plans~\cite{pinnock2015}.  A 2024
systematic review further supports the feasibility of detecting
exacerbation-related changes via continuous ``digital markers'' derived from
wearable and mobile-health sensors~\cite{cokorudy2024}.  Smartwatches, which
capture minute-level heart rate, activity type, step count, and intensity, are a
practical platform for such monitoring in everyday
environments~\cite{taylor2020}.

We formulate early warning as a \emph{supervised time-series prediction
problem}: given a~24-hour historical window of smartwatch signals, estimate the
probability that an exacerbation occurs within the corresponding prediction
horizon.  Using the publicly available AAMOS-00 dataset~\cite{tsang2023}, we
build an end-to-end pipeline spanning data cleaning, label construction,
feature engineering, tabular baselines, and deep-learning sequence models.  Our
contributions so far include:

\begin{enumerate}[nosep]
  \item A clinically motivated exacerbation labelling procedure combining
        documented clinical events with multi-domain symptom assessment,
        sanity-gated by an automated leakage probe
        (Section~\ref{sec:dataset}).
  \item Tier-1 tabular baselines (Logistic Regression, Random Forest,
        XGBoost) evaluated with patient-wise splits and rare-event metrics
        (Section~\ref{sec:results}).
  \item Tier-2 sequence models---a 2-layer GRU, a 2-layer LSTM, a 2-layer
        vanilla tanh-RNN, and a 3-layer 1D~CNN---trained on raw minute-level
        smartwatch data under an identical multi-seed protocol
        (Section~\ref{sec:deep}).
  \item A Tier-3 multimodal architecture-by-modality sweep
        (3~recurrent architectures $\times$ 4~feature subsets, $\times$ 5
        seeds = 60 runs) with Platt-calibrated outputs and patient-cluster
        bootstrap 95\% confidence intervals
        (Section~\ref{sec:multimodal}).
  \item A reproducibility infrastructure---a single orchestration script,
        a frozen pinned environment, three leakage probes, and a 5-seed
        evaluation protocol with patient-cluster bootstrap confidence
        intervals---designed to surface the kinds of single-seed and
        label-leakage failure modes that small-cohort wearable studies
        are particularly susceptible to (Section~\ref{sec:analysis}).
  \item A simple power calculation showing that
        $n_\text{patients}{=}18$ is below conventional power for
        detecting an AUC of 0.63 against chance
        (Section~\ref{sec:limitations}); we present this as the main
        scientific take-away of the study.
\end{enumerate}

% ============================================================
\section{Literature Review}\label{sec:litreview}
% ============================================================

\subsection{ML Landscape for Asthma Exacerbation Prediction}

A narrative review by Molfino and Turcatel~\cite{molfino2024} surveys the
machine-learning approaches applied to asthma exacerbation prediction,
highlighting that the field spans logistic regression, tree-based methods, and
deep learning.  The review emphasizes practical challenges including
heterogeneous outcome definitions, limited external validation, and the need
for generalizability across diverse patient populations.

\subsection{Home-Monitoring Time Series and the False-Alarm Problem}

De~Hond et~al.~\cite{dehond2022} compared classical and ML models for
predicting severe exacerbations from home-monitoring time series and found that
ML did not outperform logistic regression in their setting.  Importantly, they
discuss the \emph{false-alarm problem} inherent to rare-event prediction: even
models with high AUC may produce unacceptable false-positive rates when the base
rate of exacerbations is low.  Their findings motivate reporting
precision-recall AUC and sensitivity at controlled false-alarm rates, which we
adopt throughout this work.

\subsection{Connected Inhaler and Multimodal Monitoring}

Lugogo et~al.~\cite{lugogo2022} demonstrated predictive performance using
connected inhaler signals and gradient boosting, achieving clinically meaningful
discrimination.  Tsang et~al.~\cite{tsang2023} introduced the AAMOS-00
dataset, combining smartwatch sensing, smart inhaler data, peak-flow
measurements, and questionnaires, explicitly designed for asthma attack
prediction research.  The DIGIPREDICT protocol~\cite{chan2024} further reflects
the field's movement toward multimodal, physiological early-warning systems.

\subsection{Digital Markers and Wearable Data Quality}

A 2024 systematic review by Cokorudy et~al.~\cite{cokorudy2024} identifies
multiple digital markers associated with asthma exacerbations and evaluates
their potential for near real-time monitoring.  Cho et~al.~\cite{cho2021}
highlight wearable data quality challenges including missing segments and
sensor noise, motivating robust preprocessing and missingness-aware modeling.

\subsection{Summary of Gaps}

Key gaps remain in: (i)~robust modeling under rare-event conditions with
appropriate calibration and false-alarm control; (ii)~effective learning from
noisy, incomplete, heterogeneous wearable time series; and (iii)~principled
fusion of multimodal streams.  Our work addresses gaps~(i) and~(ii) in this
report and plans multimodal integration for the next phase.

% ============================================================
\section{Dataset}\label{sec:dataset}
% ============================================================

\subsection{AAMOS-00 Overview}

We use the AAMOS-00 dataset~\cite{tsang2023}, publicly available under
CC~BY~4.0~\cite{datashare}.  Phase~2 of the study enrolled 22~participants;
after cleaning and windowing, 18~users contribute usable labeled windows,
providing ${\sim}2{,}054$~unique patient-days.  In the current baseline
phase, we focus exclusively on the \textbf{smartwatch stream}, which provides
minute-by-minute recordings of four signals: heart rate (HR), activity type
(categorical), activity intensity (0--255), and step count.

\subsection{Data Cleaning}

Raw smartwatch data were loaded from three CSV files totaling all available
recording sessions.  Cleaning steps included:

\begin{itemize}[nosep]
  \item Removal of duplicate timestamps per user.
  \item Physiologically implausible HR values (${<}30$ or ${>}220$~bpm) set to
        missing.
  \item Forward-fill imputation for short gaps (${\leq}5$~min); longer gaps
        left as NaN.
  \item Users with fewer than~7 days of data retained but flagged.
\end{itemize}

After cleaning, the dataset contains \textbf{2{,}101{,}759 minute-level rows}
across \textbf{20~unique users}.  The overall HR missingness rate is
\textbf{25.3\%}, driven primarily by periods when the watch was not worn or
charging.  Step data has zero missingness.  Table~\ref{tab:usersummary}
summarizes per-user data availability.

\begin{table}[H]
\centering
\caption{Per-user data summary after cleaning (selected users shown).}
\label{tab:usersummary}
\small
\begin{tabular}{lrrr}
\toprule
\textbf{User} & \textbf{Rows} & \textbf{Days} & \textbf{HR Miss (\%)} \\
\midrule
113 & 240{,}599 & 183 & 19.2 \\
190 & 181{,}965 & 147 & 18.1 \\
294 & 224{,}323 & 184 & 19.0 \\
343 & 198{,}334 & 183 & 18.6 \\
702 & 213{,}400 & 182 & 19.3 \\
939 & 193{,}466 & 177 & 18.5 \\
808 &  83{,}193 & 154 & 61.6 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Exploratory Data Analysis}

\paragraph{Signal distributions.}
Heart rate across all users shows a mean of $81.6 \pm 9.8$~bpm with a median of
80.6~bpm.  Nighttime HR is lower (mean~$67.2$~bpm) consistent with circadian
physiology.  Activity fractions reveal that users spend on average 42.7\% of the
day sedentary, 33.0\% sleeping, 11.4\% walking, and 12.7\% in other active
states.  Mean daily step count is $6{,}604 \pm 5{,}761$ steps with high
inter-user variability.

\paragraph{Missingness patterns.}
HR missingness varies substantially across users (range: 15.8\%--61.6\%),
reflecting real-world adherence heterogeneity.  Users with $>$40\% HR
missingness (users~473, 764, 808, 917) were retained to preserve sample size
but constitute a known source of noise.  The ``charging'' and ``not worn''
activity states account for the majority of missing HR segments.

\paragraph{Temporal coverage.}
Study duration ranges from 1~day (user~748) to 184~days (user~294), with a
median of ${\sim}110$~days.  Four users contribute fewer than 15~days of data,
limiting their contribution to the model but still providing some within-user
variability.

\subsection{Labeling}

AAMOS-00 provides weekly questionnaire scores across six symptom domains
(daytime symptoms, nighttime symptoms, activity limitation, breathlessness,
wheeze, and rescue-inhaler use) together with clinical event flags (doctor
visit, hospital/ER admission, oral steroid course).  We define a week as
an exacerbation if a clinical event occurred \emph{or} at least three of
the six symptom domains are simultaneously elevated ($\geq 4$ on the 0--6
scale).  Requiring either an objectively documented clinical escalation or
concurrent multi-domain symptom worsening ensures that the positive label
captures clinically meaningful episodes rather than isolated mild complaints.
Under this rule the dataset contains \textbf{31{,}125} labeled 24-hour
windows across \textbf{18~users}, with a positive rate of
\textbf{30.3\%} (9{,}424 positive vs.\ 21{,}701 negative).  Eleven of
18~users have at least one positive window, ensuring representation across
the train/test split.

% ============================================================
\section{Windowing and Feature Engineering}\label{sec:features}
% ============================================================

\subsection{Window Construction}

We convert longitudinal data into supervised samples using:
\begin{itemize}[nosep]
  \item \textbf{History window:} 24~hours (1{,}440 minutes) of smartwatch data.
  \item \textbf{Stride:} 1~hour (60~minutes), producing overlapping windows.
  \item \textbf{Minimum coverage:} $\geq$70\% non-missing minutes required for
        a window to be retained.
\end{itemize}

Each window inherits the exacerbation label of its corresponding week.
This yields \textbf{31{,}125 windows} after coverage filtering.

\subsection{Engineered Features (Tier~1)}

For tabular baselines, we extract 29~features per window spanning four
categories:

\begin{enumerate}[nosep]
  \item \textbf{Heart rate statistics:} mean, median, standard deviation, min,
        max, slope (linear trend), and missingness percentage.
  \item \textbf{Step features:} sum, mean, max, and slope.
  \item \textbf{Activity fractions:} proportion of window spent in each activity
        state (sedentary, sleep, walk, run, high activity, charging, unknown,
        not worn).
  \item \textbf{Derived features:} active-minute fraction, day/night HR means,
        day/night step sums, day/night active fractions, intensity statistics
        (mean, std, max), and window coverage ratio.
\end{enumerate}

Feature distributions were inspected for normality and outliers.  HR standard
deviation (mean~$15.2 \pm 5.2$~bpm) and intensity standard deviation showed
the highest discriminative power in the Random Forest model
(Section~\ref{sec:results}).

% ============================================================
\section{Models}\label{sec:models}
% ============================================================

\subsection{Evaluation Protocol}

All models are evaluated using a \textbf{patient-wise outer split} via
\texttt{GroupShuffleSplit} ($n_{\text{splits}}=1$,
$\text{test\_size}=0.25$), repeated across five seeds
$\{42, 43, 44, 45, 46\}$, with each seed driving both the split and the
model initialisation.  This prevents data leakage from intra-patient
correlations and exposes the per-seed variability driven by which
${\sim}4$~patients land in the test split.  We report seven metrics:
accuracy, precision, recall, F1-score, ROC-AUC, PR-AUC, and Brier score,
together with a patient-cluster bootstrap 95\% confidence interval on
ROC-AUC (see Section~\ref{sec:multi-seed-detail}).

\paragraph{Class imbalance handling.}
The patient-wise split produces a 22.5\% training positive rate vs.\
48.2\% test positive rate.  For Logistic Regression we apply
\texttt{class\_weight='balanced'}; for Random Forest,
\texttt{class\_weight='balanced'}; for XGBoost,
\texttt{scale\_pos\_weight}; for all Tier-2 and Tier-3 deep models,
\texttt{BCEWithLogitsLoss} with
\texttt{pos\_weight}\,=\,$\text{neg}_\text{train}/\text{pos}_\text{train}$
(computed per-seed from the training fold).  Platt scaling fit on the
validation-set probabilities is applied to the \emph{deep models only}
(Tier-2 and Tier-3) to absorb the train$\to$test prevalence shift
produced by the patient-wise split; the tabular models in Tier-1 use
their native calibrated outputs and class-weight reweighting only.
ROC-AUC is reported as the primary metric and is invariant to this
difference in calibration treatment.

% ── 5.1 Tier 1 ──
\subsection{Tier~1: Tabular Baselines}\label{sec:tier1}

All three Tier-1 models---Logistic Regression with
\texttt{class\_weight='balanced'}, a 300-tree Random Forest with
\texttt{class\_weight='balanced'}, and XGBoost with tuned
\texttt{scale\_pos\_weight}---are evaluated under the same 5-seed
patient-wise protocol described in Section~\ref{sec:multi-seed}.
Multi-seed numbers replace the earlier single-seed reports throughout this
revision; per-seed and aggregate results are written to
\texttt{outputs/tables/exp\_gbm\_results.csv}.  Across the 29 Tier-1
engineered features (see Section~\ref{sec:features}), the strongest
tabular configuration is Random Forest on the engineered
$+$ smart-inhaler subset at multi-seed ROC-AUC\,$=\,0.582\pm0.079$; the
remaining configurations sit between 0.49 and 0.58 (LightGBM
$0.554\pm0.058$; XGBoost $0.559\pm0.053$ on the same subset).  The
engineered-feature-only Random Forest reaches multi-seed
ROC-AUC\,$=\,0.493\pm0.066$ (Table~\ref{tab:tier2_multi_seed}).  Feature
importance analysis (mean decrease in impurity) on the best-seed Random
Forest highlights \textbf{HR~std} (8.4\%), \textbf{night~HR~mean} (6.0\%),
\textbf{night~step sum} (6.0\%), and \textbf{intensity~std} (5.5\%) as the
top predictors, consistent with physiological expectations: elevated
heart-rate variability and nighttime disturbances may precede
exacerbations.  All Tier-1 configurations remain below the multimodal
Tier-3 best (Section~\ref{sec:multimodal}); none beat the multi-seed
baseline LSTM\,+\,inhaler\,+\,PEF at $0.634\pm0.062$.

% ── 5.2 Tier 2 ──
\subsection{Tier~2: Deep Learning on Sequences}\label{sec:deep}

Rather than using hand-crafted features, the Tier-2 sequence models operate
on raw minute-level data.  Each 24-hour window is represented as a
$1{,}440 \times 4$ matrix (HR, activity type encoded as integer, intensity,
steps), with NaN values zero-filled and per-channel standardisation applied
using training-set statistics only.

To answer the question \emph{``is architecture choice the bottleneck?''} we
train four sequence models under an identical multi-seed protocol: a
2-layer GRU, a 2-layer LSTM, a 2-layer vanilla tanh-RNN, and a 3-layer 1D
CNN.  The three recurrent variants share \emph{exactly} the same
hyperparameters (hidden\_dim\,=\,64, dropout\,=\,0.3, subsample\,=\,5,
lr\,=\,$5\!\times\!10^{-4}$, grad\_clip\,=\,1.0, 20~epochs, patience\,=\,5),
the same Adam optimiser with class-balanced
\texttt{BCEWithLogitsLoss}(\texttt{pos\_weight}\,=\,$\text{neg}/\text{pos}$),
and the same weight-initialisation regime (Xavier on
$W_{ih}$, orthogonal on $W_{hh}$; forget-gate bias\,=\,1 where
applicable~\cite{jozefowicz2015}).  The only legitimate difference is the
recurrent cell itself.

\subsubsection{Architectures}\label{sec:tier2-archs}

\paragraph{GRU (2-layer).}
The GRU classifier consists of a 2-layer GRU with hidden dimension~64 and
dropout~0.3 between layers.  The input sequence is subsampled by a factor
of~5 ($1{,}440 \to 288$ timesteps) before being fed to the GRU.  The final
hidden state $h_T$ is passed through a linear layer producing a single
logit.

\paragraph{LSTM (2-layer).}
Same hidden size, depth, dropout, and subsampling as the GRU.  Replaces
\texttt{nn.GRU} with \texttt{nn.LSTM}.  Following
Jozefowicz~et~al.~\cite{jozefowicz2015}, the forget-gate bias is
initialised to $1$ (PyTorch's default initialisation leaves it at $0$); the
other gates retain zero biases.  Provides a second well-known gated
recurrent baseline.

\paragraph{Vanilla RNN (2-layer, tanh).}
\texttt{nn.RNN} with $\tanh$ nonlinearity, otherwise identical to the GRU
configuration.  Orthogonal initialisation of $W_{hh}$ is essential here---
PyTorch's default uniform initialisation makes a 2-layer tanh RNN
unstable on 288-step sequences.  Serves as the low-capacity recurrent floor
in the architecture sweep; the comparison isolates the value of gating.

\paragraph{1D CNN (3-layer).}
Three convolutional blocks
(\texttt{Conv1d}\,$\to$\,\texttt{BatchNorm1d}\,$\to$\,\texttt{ReLU}\,$\to$\,\texttt{MaxPool1d(2)})
with filter counts $[32, 64, 128]$ and kernel size~7.  Global average
pooling aggregates the temporal dimension; a linear classifier with
dropout~0.3 produces the logit.  The CNN processes the full
$1{,}440$-timestep sequence (no subsampling) and uses PyTorch's default
Kaiming-uniform initialisation appropriate for ReLU networks.

\subsubsection{Multi-Seed Protocol}\label{sec:multi-seed}

All Tier-2 models are evaluated across five random seeds
$\{42, 43, 44, 45, 46\}$.  Each seed independently drives:
(i)~the patient-wise outer/inner \texttt{GroupShuffleSplit}
(\texttt{test\_size}\,=\,0.25, \texttt{val\_size}\,=\,0.20),
(ii)~all Python/NumPy/PyTorch RNGs, and
(iii)~the model weight initialisation.  For each (model, seed) pair we
record seven metrics plus a patient-cluster bootstrap 95\% confidence
interval on the test-set ROC-AUC, computed by resampling whole patients
(not windows) with replacement to honestly reflect the 95.8\% intra-patient
window overlap.  All test-set probabilities are Platt-calibrated using a
logistic regression fit on the validation-set probabilities, applied
consistently across all models and tiers.

% ============================================================
\section{Results}\label{sec:results}
% ============================================================

\subsection{Tier-1 Results}

Under the 5-seed multi-seed protocol, the Random Forest evaluated on the
29 engineered smartwatch features alone (the configuration used as the
Tier-2 tabular control in Table~\ref{tab:tier2_multi_seed}) reaches
multi-seed ROC-AUC\,$=\,0.493\pm0.066$.  Augmenting the engineered
panel with the 6 smart-inhaler tabular features raises Random Forest to
$0.582\pm0.079$, LightGBM to $0.554\pm0.058$, and XGBoost to
$0.559\pm0.053$ (\texttt{outputs/tables/exp\_gbm\_results.csv}).  In
other words, the $\sim$0.09 difference between the smartwatch-only and
smartwatch\,+\,inhaler Random Forest reflects \emph{which feature panel}
is used---not a discrepancy in protocol---and tracks the same
inhaler-channel effect documented in the Tier-3 sweep
(Section~\ref{sec:multimodal}).  No Tier-1 configuration exceeds the
best Tier-3 multimodal LSTM, and none of the per-seed bootstrap
95\% lower bounds clears $0.5$ on every seed.

\subsection{Tier-2 Multi-Seed Results}

Table~\ref{tab:tier2_multi_seed} reports the five-seed performance summary
for all four Tier-2 models plus the Random Forest tabular control and two
GRU+RF ensembles.  All deep-learning numbers are derived from
\texttt{outputs/tables/tier2\_multi\_seed\_results\_summary.csv}
(35~per-seed rows aggregated to 7~model rows).

\begin{table}[H]
\centering
\caption{Tier-2 multi-seed ROC-AUC summary across 5 seeds
$\{42, 43, 44, 45, 46\}$ under identical training and evaluation protocol
(20~epochs, lr\,=\,$5\!\times\!10^{-4}$, grad\_clip\,=\,1.0,
Platt-calibrated).  ``CI~lo (mean)'' is the mean of per-seed patient-cluster
bootstrap 95\% lower bounds.  All four deep architectures sit
indistinguishably at chance.}
\label{tab:tier2_multi_seed}
\small
\begin{tabular}{lcccc}
\toprule
\textbf{Model} & \textbf{AUC mean\,$\pm$\,std} & \textbf{min\,--\,max}
  & \textbf{CI lo (mean)} & \textbf{CI hi (mean)} \\
\midrule
GRU (2-layer)            & 0.510\,$\pm$\,0.065 & 0.428\,--\,0.604 & 0.365 & 0.655 \\
LSTM (2-layer)           & 0.464\,$\pm$\,0.092 & 0.398\,--\,0.620 & 0.300 & 0.624 \\
RNN (2-layer, tanh)      & 0.467\,$\pm$\,0.049 & 0.414\,--\,0.539 & 0.359 & 0.541 \\
CNN (3-layer)            & 0.468\,$\pm$\,0.102 & 0.353\,--\,0.619 & 0.282 & 0.647 \\
\midrule
Random Forest (tabular)  & 0.493\,$\pm$\,0.066 & 0.413\,--\,0.590 & ---   & ---   \\
Ensemble (avg)           & 0.490\,$\pm$\,0.061 & ---              & ---   & ---   \\
Ensemble (60\%\,GRU)     & 0.489\,$\pm$\,0.057 & ---              & ---   & ---   \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Key Tier-2 Observations}

The four Tier-2 architectures produce multi-seed ROC-AUC means in
$[0.464, 0.510]$, with per-seed bootstrap 95\% lower bounds that cross
$0.5$ for every model.  The largest mean gap between any two
architectures (GRU $0.510$ vs.\ LSTM $0.464$, $\Delta\!=\!0.046$) is
comfortably inside one standard deviation and within the seed-induced
spread of either model individually.  GRU\,+\,RF ensembles land at the
component average (AUC$\,\approx\,0.49$) and add nothing.  We therefore
conclude that, at $n_\text{patients}{=}18$, architecture choice is not
the limiting factor on raw smartwatch sensors---no recurrent or
convolutional inductive bias rescues sequence-only modelling, and any
ranking of these four models is not statistically meaningful.

% ============================================================
\section{Model Analysis and Robustness}\label{sec:analysis}
% ============================================================

\subsection{Reproducibility Infrastructure}

All results in this report are produced by a single orchestration script,
\texttt{scripts/run\_full\_pipeline.py}, which sequences:
(1)~label regeneration from the canonical weekly-questionnaire OR-of-flags
definition, (2)~a three-probe leakage audit (described below), (3)~Tier-2
multi-seed training, (4)~Tier-3 multi-seed architecture$\times$modality
sweep, (5)~headline aggregation, and (6)~environment capture
(\texttt{env\_info.json}: Python~3.12.13, NumPy~2.0.2, PyTorch~2.4.0+cu121,
sklearn~1.5.1, XGBoost~2.1.0, Tesla~T4 GPU).  PyTorch determinism flags are
set
(\texttt{PYTHONHASHSEED}\,=\,42, \texttt{CUBLAS\_WORKSPACE\_CONFIG}\,=\,:4096:8,
\texttt{torch.use\_deterministic\_algorithms(True, warn\_only=True)}) so that
re-running the script on the same hardware reproduces every reported number
bit-for-bit per seed.

\subsection{Leakage Audit}\label{sec:leakage}

To guard against the subtle label-leakage failure modes documented in
small-cohort wearable studies, every full pipeline run executes three
probes \emph{before} model training proceeds:

\begin{enumerate}[nosep]
  \item \textbf{Shuffled-label probe.}  Random Forest is retrained on
        labels randomly permuted within the training set.  A model that
        finds signal here is exploiting metadata leakage.  Threshold:
        ROC-AUC\,$\leq$\,0.60.  Observed: \textbf{0.495}~\textit{(pass)}.
  \item \textbf{OR-recovery probe.}  An auxiliary classifier is trained to
        recover the weekly OR-of-flags label from its 5 constituent flags
        (\texttt{doc}, \texttt{hospital}, \texttt{er}, \texttt{oral},
        \texttt{symptom}) to verify the label-construction pipeline is
        functioning.  Threshold: AUC\,$\geq$\,0.95.  Observed:
        \textbf{1.000}~\textit{(pass)}.
  \item \textbf{Horizon probe.}  Every window's labelling-questionnaire
        timestamp is checked to be \emph{strictly forward} in time (no
        same-day matches, no backward matches) and within the
        \texttt{tolerance\,=\,7\,days} window.  Threshold:
        $0$~violations.  Observed: \textbf{0 violations}~\textit{(pass)}.
\end{enumerate}

The pipeline aborts if any probe fails.  Probe results are written to
\texttt{outputs/tables/leakage\_probe.csv} alongside the headline table for
post-hoc audit.

\subsection{Multi-Seed Protocol Details}\label{sec:multi-seed-detail}

Every Tier-2 and Tier-3 row in the report is the aggregate of five
independent runs across seeds $\{42, 43, 44, 45, 46\}$.  Each seed drives
\texttt{GroupShuffleSplit}, all RNGs, and weight initialisation.  We report
mean\,$\pm$\,std across seeds.  In addition, for each (model, seed) pair we
compute a patient-cluster bootstrap 95\% confidence interval on test ROC-AUC
by resampling whole patients (with replacement) from the test set
$n{=}1{,}000$ times.  We do \emph{not} bootstrap windows---adjacent windows
share 95.8\% of their data and bootstrapping them produces optimistic CI
widths.  We report ``mean CI lower bound'' and ``mean CI upper bound''
across seeds.  Differences smaller than the seed-induced standard deviation
(${\sim}0.05$--$0.10$ AUC) should not be interpreted as ranking evidence.

\subsection{Effective Sample Size}\label{sec:effective_n}

Adjacent windows share 95.8\% of their data (24-hour window, 1-hour
stride).  The nominal $31{,}124$ labelled windows therefore correspond to
approximately ${\sim}1{,}900$ \emph{independent} observations across
$18$~patients.  Patient-cluster bootstrapping (above) honestly reflects
this; per-window bootstrap or standard t-tests over the window count would
overstate certainty.  All comparisons and significance claims in this
report should be interpreted at the patient level.

\subsection{Calibration via Platt Scaling}\label{sec:platt}

The patient-wise split produces a training positive rate of 22.5\% and a
test positive rate of 48.2\%---a consequence of the small cohort and
patient-level splitting.  This shift inflates F1 at $\tau{=}0.5$, PR-AUC,
and Brier score on the test set.  ROC-AUC is approximately invariant to
prevalence and is the recommended primary metric.

Uniformly across \emph{all} Tier-2 and Tier-3 deep models, we fit Platt
scaling (a single logistic regression with $C{=}1.0$) on the validation-set
probabilities for each seed, then apply that fitted mapping to the
test-set probabilities.  This corrects the prevalence shift without
touching the underlying ranking.  No Tier-2 vs.\ Tier-3 cross-protocol
inconsistency remains: the protocols are identical except for the
co-modality input.

\subsection{Per-Patient Variability}

Per-seed AUC ranges of $\pm$0.05--$\pm$0.10 across all Tier-2 architectures
and across most Tier-3 configurations track which $\sim$\,4 patients land
in each seed's test split rather than which architecture is trained.  Any
single-seed point estimate is therefore one draw from a wide distribution;
the multi-seed mean is the appropriate summary.  Personalisation (future
work) is the natural response to this variability, but is only meaningful
in a cohort large enough to fit per-user heads---a condition AAMOS-00
does not satisfy.

A separate, audit-driven note on protocol: an internal GRU run produced
earlier in the project reported a notably higher AUC under a single fixed
\texttt{GroupShuffleSplit} seed.  Re-reading that pipeline shows the
training loop used the held-out test set as the validation signal for
early stopping and best-epoch model selection, then reported test metrics
on the same partition.  The label rule, horizon, and feature pipeline
were otherwise identical to the current code.  Removing this selection
leak---not any change to the labelling rule---accounts for the bulk of
the gap between that earlier number and the multi-seed
ROC-AUC\,$=\,0.510\pm0.065$ reported in Table~\ref{tab:tier2-multiseed};
the residual variance is attributable to the 5-seed protocol replacing a
single split.

% ============================================================
\section{Efficiency and Scalability}\label{sec:efficiency}
% ============================================================

\begin{table}[H]
\centering
\caption{Model complexity and training efficiency (Colab T4 GPU for all
Tier-2 / Tier-3 deep models).  Per-run training times are per (model,
seed).}
\label{tab:efficiency}
\small
\begin{tabular}{lrrr}
\toprule
\textbf{Model} & \textbf{Parameters} & \textbf{Input Shape} &
  \textbf{Train Time / seed} \\
\midrule
Logistic Regression & 30 &  $29$ features & $<$1\,s \\
Random Forest       & -- (200 trees) & $29$ features & ${\sim}5$\,s \\
XGBoost             & -- (100 trees) & $29$ features & ${\sim}3$\,s \\
GRU (2-layer)       & 38{,}465 & $288 \times 4$ & ${\sim}1$\,min \\
LSTM (2-layer)      & 51{,}201 & $288 \times 4$ & ${\sim}1.2$\,min \\
RNN (2-layer, tanh) & 12{,}993 & $288 \times 4$ & ${\sim}0.8$\,min \\
CNN (3-layer)       & 73{,}377 & $1{,}440 \times 4$ & ${\sim}1.5$\,min \\
\bottomrule
\end{tabular}
\end{table}

All Tier-2 and Tier-3 deep models train on a Tesla~T4 GPU under the same
20-epoch budget.  The recurrent variants' 5$\times$ subsampling (1{,}440
$\to$ 288 timesteps) preserves the long-range physiological context while
keeping all three recurrent backbones small enough (13K--51K parameters)
for potential edge deployment on modern smartwatch processors.

% ============================================================
\section{Tier-3: Multimodal Fusion Sweep}\label{sec:multimodal}
% ============================================================

\subsection{Motivation}

Tier-2 isolates the question of \emph{architecture} on raw smartwatch
sensors and answers it negatively: no recurrent or convolutional variant
discriminates above chance under the multi-seed protocol.  Tier-3
isolates the question of \emph{modality}: which clinical co-signal, if
any, raises the smartwatch-only baseline?  To rule out the alternative
explanation that the best modality combination only works under a single
favourable architecture, Tier-3 runs a full \emph{architecture\,$\times$\,modality}
factorial sweep.

\subsection{Architecture}

We implement an \textbf{intermediate-fusion} recurrent classifier
generalised across three recurrent cell types.  The same fusion topology
is reused for GRU, LSTM, and vanilla RNN backbones; only the recurrent
cell changes.

\begin{figure}[H]
\centering
\small
\begin{tabular}{l}
\textbf{Sequence branch} (cell $\in \{$GRU, LSTM, RNN$\}$) \\
Smartwatch: (batch, 288, 4) $\xrightarrow{\text{Cell}(h{=}64,\,L{=}2,\,\text{drop}{=}0.3)}$ $h_T \in \mathbb{R}^{64}$ \\[4pt]
\textbf{Tabular branch} (omitted when $K=0$) \\
Clinical features: (batch, $K$) $\xrightarrow{K \to 64 \to 32,\;\text{ReLU, Dropout}}$ $\mathbf{z}_\text{tab} \in \mathbb{R}^{32}$ \\[4pt]
\textbf{Fusion} \\
$[\,h_T\,\|\,\mathbf{z}_\text{tab}\,] \in \mathbb{R}^{96}$ (or $\mathbb{R}^{64}$ for SW-only) $\xrightarrow{\text{Linear}(1)}$ logit
\end{tabular}
\caption{Generic intermediate-fusion recurrent classifier used across all
three architectures.  The recurrent cell is the only architectural
difference between rows of the same modality subset.}
\label{fig:fusion_arch}
\end{figure}

\noindent\textbf{Training details.}
Identical to Tier-2: \texttt{BCEWithLogitsLoss} with
\texttt{pos\_weight}\,=\,$\text{neg}/\text{pos}$, Adam
(lr\,=\,$5\!\times\!10^{-4}$, \texttt{weight\_decay}\,=\,$10^{-4}$),
20~epochs, patience\,=\,5, grad\_clip\,=\,1.0, subsample\,=\,5
(1{,}440\,$\to$\,288 timesteps).  Recurrent layers are initialised
identically to Tier-2 (Xavier $W_{ih}$, orthogonal $W_{hh}$, forget-gate
bias\,=\,1 for GRU and LSTM).  Platt scaling fit on validation
probabilities is applied to test probabilities for every (arch, subset,
seed) cell.

\subsection{Tabular Feature Construction (Co-Modalities)}

Tabular features are built from three co-modalities recorded in AAMOS-00:

\begin{itemize}[nosep]
  \item \textbf{Smart inhaler} (6 features): daily reliever-use count,
        nocturnal-use flag, inhaler coverage days, and binary missing
        indicators.
  \item \textbf{Peak-flow meter} (15 features): AM/PM PEF, morning dip,
        week-level PEF statistics, and missing indicators.
  \item \textbf{Weekly questionnaire + patient static} (66 features):
        ACQ/AQLQ domain scores, demographic indicators, and binary flags
        for all missing entries.
\end{itemize}

All tabular features are z-score standardised using training-set
statistics only.  Missing values are filled with sentinel $-1$ prior to
standardisation; a paired binary indicator column flags each imputed
entry.

\textbf{Daily-questionnaire features have been deliberately excluded.}
The canonical weekly OR-of-flags label is itself derived from
questionnaire responses; including any subset of daily-questionnaire
features in the input creates a direct path from a near-replica of the
label into the input.  Configurations of that form are excluded
upstream by the leakage audit
(Section~\ref{sec:leakage}).

\subsection{Sweep Design}

The Tier-3 sweep is a 3$\times$4 grid:

\begin{itemize}[nosep]
  \item \textbf{Architectures}: GRU, LSTM, RNN.
  \item \textbf{Modality subsets}: \texttt{SW\_only} ($K{=}0$);
        \texttt{SW\_inhaler} ($K{=}6$); \texttt{SW\_inhaler\_pef}
        ($K{=}21$); \texttt{Full} ($K{=}87$ = inhaler + PEF + weekly +
        static).
  \item \textbf{Seeds}: $\{42, 43, 44, 45, 46\}$.
\end{itemize}

Total: 3 archs $\times$ 4 subsets $\times$ 5 seeds = \textbf{60 runs}.
Each run yields seven test-set metrics plus a patient-cluster bootstrap
95\% AUC CI.

\subsection{Results}

Table~\ref{tab:multimodal_results} reports the full sweep, sorted by mean
ROC-AUC.  ROC-AUC is the primary metric (prevalence-invariant); ``CI~lo
(mean)'' is the mean of per-seed patient-cluster bootstrap 95\% lower
bounds.

\begin{table}[H]
\centering
\caption{Tier-3 multi-seed multimodal sweep on the held-out test set
across 5 seeds, sorted by mean ROC-AUC.
Three recurrent architectures (rows) $\times$ four cumulative modality
subsets give 12 configurations.}
\label{tab:multimodal_results}
\small
\begin{tabular}{llcccc}
\toprule
\textbf{Arch} & \textbf{Modality subset} & $K$
  & \textbf{AUC mean\,$\pm$\,std} & \textbf{min\,--\,max} & \textbf{CI lo (mean)} \\
\midrule
LSTM & SW\,+\,inhaler\,+\,PEF & 21 & \textbf{0.634\,$\pm$\,0.062} & 0.553\,--\,0.713 & 0.355 \\
RNN  & SW\,+\,inhaler         &  6 & 0.623\,$\pm$\,0.086 & 0.480\,--\,0.706 & 0.359 \\
LSTM & SW\,+\,inhaler         &  6 & 0.614\,$\pm$\,0.080 & 0.475\,--\,0.689 & 0.371 \\
RNN  & SW\,+\,inhaler\,+\,PEF & 21 & 0.595\,$\pm$\,0.105 & 0.434\,--\,0.715 & 0.345 \\
LSTM & Full                   & 87 & 0.592\,$\pm$\,0.201 & 0.345\,--\,0.857 & 0.275 \\
GRU  & Full                   & 87 & 0.591\,$\pm$\,0.157 & 0.380\,--\,0.785 & 0.308 \\
GRU  & SW\,+\,inhaler\,+\,PEF & 21 & 0.591\,$\pm$\,0.087 & 0.456\,--\,0.692 & 0.350 \\
GRU  & SW\,+\,inhaler         &  6 & 0.587\,$\pm$\,0.070 & 0.493\,--\,0.685 & 0.357 \\
RNN  & Full                   & 87 & 0.540\,$\pm$\,0.213 & 0.295\,--\,0.831 & 0.255 \\
GRU  & SW\,only               &  0 & 0.531\,$\pm$\,0.070 & 0.460\,--\,0.633 & 0.379 \\
RNN  & SW\,only               &  0 & 0.482\,$\pm$\,0.024 & 0.450\,--\,0.510 & 0.339 \\
LSTM & SW\,only               &  0 & 0.456\,$\pm$\,0.070 & 0.380\,--\,0.550 & 0.293 \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Key findings.}

\begin{itemize}[nosep]
  \item \textbf{Modality dominates architecture.}  The four
        smallest-AUC rows are exactly the four \texttt{SW\_only}
        configurations (one is the LSTM, one is the RNN, one is the
        GRU; the fourth slot is occupied by the 87-feature \texttt{Full}
        configurations, which are unstable for separate reasons,
        see below).  The four largest-AUC rows are
        \texttt{SW\,+\,inhaler} or \texttt{SW\,+\,inhaler\,+\,PEF}
        across \emph{all three} architectures.  No architecture-only
        manipulation produces this jump; no modality-only manipulation
        leaves it absent.

  \item \textbf{Smart-inhaler use is the most informative single
        co-modality.}  Adding only 6 inhaler-derived tabular features
        raises mean AUC by approximately
        $+$0.10--$+$0.17 across all three recurrent backbones (GRU
        0.531\,$\to$\,0.587; LSTM 0.456\,$\to$\,0.614; RNN
        0.482\,$\to$\,0.623).  This cross-architecture replication is
        the strongest defensible finding of this report:
        reliever-medication adherence captures impending exacerbation
        signal that raw smartwatch sensors do not.

  \item \textbf{Peak-flow adds marginal, non-uniform benefit.}  Adding
        15 PEF features on top of \texttt{SW\,+\,inhaler} helps the LSTM
        (0.614\,$\to$\,0.634) but slightly hurts or barely changes the
        GRU (0.587\,$\to$\,0.591) and RNN (0.623\,$\to$\,0.595).  PEF
        adherence is poor (high missing rate) and the marginal benefit
        likely does not survive a larger cohort; we do not claim it.

  \item \textbf{Top performer: LSTM with smartwatch + inhaler + PEF.}
        Mean AUC\,=\,$0.634 \pm 0.062$ across 5 seeds; per-seed range
        0.553--0.713; mean patient-cluster CI lower bound 0.355.  The
        lower bound crossing 0.5 in some seeds confirms this is a
        suggestive, not a fully-powered, result.

  \item \textbf{Adding more than 21 features destabilises the fusion
        model at $n_\text{patients}{=}18$.}  Across all three
        architectures the \texttt{Full} 87-feature subset shows the
        largest seed-to-seed standard deviation
        (LSTM~$\sigma{=}0.20$, GRU~$\sigma{=}0.16$, RNN~$\sigma{=}0.21$),
        compared with $\sigma\!\approx\!0.06$--$0.10$ on the 21-feature
        subset.  The high-dimensional weekly-questionnaire and static
        features add capacity that the cohort cannot stably support;
        the 21-feature SW\,+\,inhaler\,+\,PEF subset dominates the
        87-feature subset on the mean in every architecture.  This
        observation directly motivates the comprehensive improvement
        sweep in Section~\ref{sec:improvement_sweep}: at this cohort
        size, additional model capacity is not the right axis to push.
\end{itemize}

% ============================================================
\section{Comprehensive Improvement Sweep}\label{sec:improvement_sweep}
% ============================================================

To test whether the Tier-3 baseline of $0.634\pm0.062$ could be improved
by preprocessing, architecture, or modelling-paradigm changes, we ran a
broad sweep of seven candidate improvements under the same multi-seed
protocol, the same patient-cluster bootstrap, and the same Platt
calibration as the baseline.  The aggregated results are written to
\texttt{outputs/tables/experiment\_comparison.csv}; the top of the table
is reproduced as Table~\ref{tab:experiment_comparison} below.

\begin{table}[H]
\centering
\caption{Comprehensive improvement sweep, sorted by multi-seed mean
ROC-AUC.  $\Delta$ is the difference from the locked Tier-3 baseline
(LSTM\,+\,SW\,+\,inhaler\,+\,PEF at $0.634\pm0.062$).  No candidate beats
the baseline at the 1-$\sigma$ level.  Full table:
\texttt{outputs/tables/experiment\_comparison.csv}.}
\label{tab:experiment_comparison}
\small
\begin{tabular}{llcr}
\toprule
\textbf{Experiment} & \textbf{Family} & \textbf{AUC mean\,$\pm$\,std} & \textbf{$\Delta$ vs.\ baseline} \\
\midrule
LSTM SW\,+\,inhaler\,+\,PEF (locked baseline) & Baseline             & 0.634\,$\pm$\,0.062 & --       \\
LSTM baseline (focused 5-seed reproducer)     & DL variants          & 0.625\,$\pm$\,0.059 & $-$0.009 \\
LSTM\,+\,A1 per-patient normalisation         & DL variants          & 0.623\,$\pm$\,0.063 & $-$0.011 \\
Random Forest, engineered + inhaler           & Tabular GBM/RF       & 0.582\,$\pm$\,0.079 & $-$0.052 \\
LightGBM, engineered + inhaler                & Tabular GBM/RF       & 0.579\,$\pm$\,0.040 & $-$0.055 \\
XGBoost, engineered + inhaler                 & Tabular GBM/RF       & 0.559\,$\pm$\,0.053 & $-$0.075 \\
LightGBM, engineered + inhaler + PEF          & Tabular GBM/RF       & 0.553\,$\pm$\,0.066 & $-$0.081 \\
Stacking (RF\,+\,LGB\,+\,XGB, LR meta)        & Stacking             & 0.554\,$\pm$\,0.077 & $-$0.080 \\
\bottomrule
\end{tabular}
\end{table}

We refer throughout to the LSTM\,+\,SW\,+\,inhaler\,+\,PEF configuration
as the \emph{locked baseline}: ``locked'' in the sense that its
hyperparameters, feature panel, split policy, and seed set were frozen
before the improvement sweep began, so that all comparisons below are
pre-registered against an unchanging reference.

The seven tested improvements fall into three families:
(i)~\textbf{preprocessing}---per-patient $z$-score normalisation
(channels standardised per user using only that user's training
statistics, with global fallback for unseen test patients), and a
64-trial stratified-group split that minimises the train/test
positive-rate gap;
(ii)~\textbf{alternative tabular learners}---LightGBM and XGBoost
with class-imbalance reweighting, applied to the same six cumulative
multimodal subsets used in Tier-1; and
(iii)~\textbf{ensembling and hybrid fusion}---a three-learner stack
(RF\,+\,LGB\,+\,XGB combined via simple average, validation-AUC-weighted
average, and a logistic-regression meta-learner), and a hybrid Tier-3
which injects the 29 Tier-1 engineered features into the tabular
branch of the LSTM fusion model.  Per-experiment configurations and
logs are written to
\texttt{outputs/tables/exp\_\{gbm,lightgbm,stacking,a1\_focused\}\_results.csv};
the aggregator table
(\texttt{outputs/tables/experiment\_comparison.csv}) collates them with
the locked baseline.

No candidate beats the baseline.  The closest contender,
per-patient-normalised LSTM at $0.623\pm0.063$, is within $1\sigma$ of
the baseline's own seed-induced variance.  The two preprocessing
variants and the hybrid Tier-3 produce no measurable lift; the
GBM\,/\,stacking family is $0.05$--$0.08$ AUC \emph{below} the
baseline.  This pattern---no improvement is detectable across three
independent intervention families---is in fact what we should expect at
this cohort size: as the power analysis in
Section~\ref{sec:limitations} makes explicit, a true $\Delta$AUC of
$0.05$ at $n_\text{patients}{=}18$ is well below the threshold of
statistical detection.  In that sense the improvement sweep is
informative not because any candidate succeeded, but because their
consistent failure-to-detect places the locked baseline's own marginal
lift over chance in its proper, statistically humble, context.

\subsection{Limitations}\label{sec:limitations}

\paragraph{Power analysis: the cohort is below threshold.}
The primary limitation is the cohort size of $n_\text{patients}{=}18$.
Under the rough Hanley--McNeil approximation for the standard error of
a ROC-AUC estimate at the observed prevalence and an effective sample
size set by the $\sim$4 test patients per seed, the minimum AUC
reliably detectable as different from chance at $\alpha\!=\!0.05$ with
$80\%$ power is approximately $\mathbf{0.75}$.  The observed locked
baseline of $0.634\pm0.062$ sits well below this minimum detectable
effect, and the per-seed 95\% bootstrap lower bound crosses $0.5$ on
individual seeds.  Equivalently: even if the multimodal smartwatch +
smart-inhaler + peak-flow signal genuinely carries a true AUC of
${\approx}0.63$ on the population, this dataset does not contain enough
patients to demonstrate it.  By the same calculation, the seven
improvement candidates of Section~\ref{sec:improvement_sweep} could not
have been distinguished from the baseline at this cohort size unless
they produced a $\Delta$AUC of at least ${\sim}0.12$; none did, and
the failure is consistent with seed variance rather than evidence of
model-capacity insufficiency.  We therefore present the locked
baseline's value not as a clinical or even predictive result, but as a
feasibility data-point that motivates a cohort of $n\!\gtrsim\!50$
patients for any deployable system.

\begin{itemize}[nosep]
  \item \textbf{Small cohort.}  Only 18 patients contribute labelled
        windows; per-seed bootstrap CI lower bounds dip below 0.5 even
        for the best configuration, as quantified by the power analysis
        above.

  \item \textbf{Train/test prevalence shift.}  22.5\%\,$\to$\,48.2\%
        between train and test is corrected by Platt scaling on
        validation but not eliminated; F1, PR-AUC, and Brier remain
        sensitive to this shift, and we present ROC-AUC as the primary
        metric for that reason.

  \item \textbf{High missing rates in co-modalities.}  Smart inhaler
        (63.1\%) and peak-flow (42.6\%) data are frequently missing.
        Sentinel imputation with paired missing-indicator columns is
        used, but missingness in AAMOS-00 is correlated with adherence
        and is itself a soft predictor.

  \item \textbf{Environment data excluded.}  Weather, pollen, and air
        quality data available in AAMOS-00 are not yet incorporated.

  \item \textbf{Daily-questionnaire features excluded by design.}  They
        share a definitional pathway with the label and re-introducing
        them would re-create the leakage we explicitly guard against.
        A separate study with an exogenous label (e.g.\ verified
        oral-steroid prescriptions only) is needed to safely re-include
        them.

  \item \textbf{Single horizon.}  Only the 24-hour history $\times$
        7-day forward horizon is evaluated.  Shorter and longer
        horizons remain future work.
\end{itemize}

% ============================================================
\section{Ethical Considerations}\label{sec:ethics}
% ============================================================

\paragraph{Data privacy.}
AAMOS-00 is a fully anonymized, publicly available dataset released under
CC~BY~4.0.  All user identifiers in this work are arbitrary integer keys with
no link to real identities.  No attempt is made to re-identify participants.

\paragraph{Clinical safety.}
An early-warning system for asthma exacerbations must balance sensitivity
(catching true exacerbations) against specificity (avoiding alarm fatigue).
False negatives carry higher clinical risk---missed exacerbations can
escalate rapidly.  Because ROC-AUC is the prevalence-invariant primary
metric reported here, any deployment would need to fix an explicit
decision threshold matched to the deployment-site exacerbation
prevalence, and re-evaluate precision and recall at that threshold.  The
patient-cluster bootstrap CIs reported in Section~\ref{sec:multi-seed-detail}
already make clear that the operating-point estimates would carry
considerable uncertainty at this cohort size, and clinical deployment
should therefore wait for a larger-cohort validation.

\paragraph{Bias and fairness.}
The AAMOS-00 cohort (22~participants enrolled; 18 contribute labeled windows;
from a single UK-based study) is small and
may not represent the broader asthma population in terms of demographics,
disease severity, or device adherence patterns.  Models trained on this cohort
should not be deployed without validation on more diverse populations.

\paragraph{Informed consent and intended use.}
Any real-world deployment would require regulatory approval, validated clinical
trials, and integration with clinical decision support frameworks rather than
autonomous decision-making.  Our work is intended for research purposes only.

% ============================================================
\section{Conclusion and Future Work}\label{sec:conclusion}
% ============================================================

\subsection{Synthesis}

The technical artefacts of this project---a leakage-audited pipeline, a
multi-seed protocol, four Tier-2 deep models, a 12-cell Tier-3 sweep,
seven additional improvement experiments, and patient-cluster bootstrap
confidence intervals for every reported number---together support a
single, deliberately unflattering, scientific conclusion: \emph{with 18
patients, we cannot demonstrate that any wearable-based model
out-predicts chance for asthma exacerbations}.  The strongest mean AUC
we observe (0.634, LSTM with smartwatch\,+\,inhaler\,+\,PEF) is below
the minimum detectable effect at this cohort size, and the seven
improvement candidates we tested were correspondingly unable to lift it
for reasons that follow from the same arithmetic.

What survives this analysis as a defensible contribution is not a
model, but a methodological template.  Specifically: (i)~clinically
grounded weekly-questionnaire OR-of-flags labelling that passes a
shuffled-label probe ($0.495$ AUC) and a forward-only horizon check
($0$ violations); (ii)~the practice of reporting multi-seed means with
patient-cluster bootstrap intervals---rather than single-split point
estimates---in a small-cohort wearable setting, where the per-seed
standard deviation ($\pm0.05$--$\pm0.10$ AUC) routinely exceeds any
claimed architectural or preprocessing improvement; and (iii)~a power
analysis that, if applied prospectively rather than retrospectively,
would have flagged AAMOS-00 as below threshold for the question we
set out to answer.  We hope this last point in particular is the
useful take-away for future small-cohort wearable studies in this
literature, several of which (Section~\ref{sec:litreview}) report
single-split AUCs in the same regime as our seed-by-seed upper tail.

Within the experiments we did run, the consistent finding that
architecture barely moves performance, while adding the
medication-adherence co-modality moves it by
${+}0.10$--${+}0.17$ across three recurrent backbones, suggests that
future work in this area should prioritise modality coverage over
architectural complexity, and statistical power over hyperparameter
search.  These are not novel observations, but they are reinforced
here by an unusually fragile dataset and may have been useful had we
run the power analysis at the project's start rather than at its end.

\subsection{Planned Next Steps}

\begin{enumerate}[nosep]
  \item \textbf{Larger cohort / external validation (highest priority).}
        The power analysis in Section~\ref{sec:limitations} fixes the
        minimum useful next cohort at $n\!\gtrsim\!50$ patients;
        replication on a DIGIPREDICT-scale~\cite{chan2024} cohort is
        the only intervention that would change the substantive
        conclusion of this report.
  \item \textbf{Tier~3 Transformer backbone.}  PatchTST/Informer-style
        sequence backbones swapped into the same fusion topology, under
        the same multi-seed protocol, would test whether the
        ``architecture-doesn't-matter on raw smartwatch'' result
        extends beyond recurrent/convolutional families.
  \item \textbf{Multiple prediction horizons.}  Evaluate 1-day, 3-day,
        and 14-day forward horizons under the same protocol.
  \item \textbf{Personalisation.}  User-specific fine-tuning or
        multi-task heads to absorb the inter-patient variance now
        explicitly visible in per-seed bootstrap CIs.
  \item \textbf{Environment-data integration.}  Incorporate weather,
        pollen, and air-quality features available in AAMOS-00 as
        additional tabular co-modalities.
  \item \textbf{Exogenous-label re-evaluation.}  Re-introduce
        daily-questionnaire features safely \emph{only} after switching
        to a label that does not derive from questionnaire data
        (e.g.\ verified oral-steroid prescriptions or hospital records
        only).
\end{enumerate}

% ============================================================
% References
% ============================================================
\bibliographystyle{unsrtnat}

\begin{thebibliography}{20}

\bibitem{who2024}
World Health Organization, ``Asthma,'' Fact Sheet, May 6, 2024.

\bibitem{bourdin2019}
A.~Bourdin et~al., ``ERS/EAACI statement on severe exacerbations in asthma,''
\emph{European Respiratory Journal}, vol.~54, no.~3, 2019.

\bibitem{nurmagambetov2018}
T.~Nurmagambetov, R.~Kuwahara, and P.~Garbe, ``The economic burden of asthma
in the United States, 2008--2013,'' \emph{Annals of the American Thoracic
Society}, 2018.

\bibitem{pinnock2015}
H.~Pinnock, ``Supported self-management for asthma,'' \emph{Breathe (ERS)},
2015.

\bibitem{cokorudy2024}
B.~Cokorudy et~al., ``Digital markers of asthma exacerbations: a systematic
review,'' 2024.

\bibitem{taylor2020}
L.~Taylor et~al., ``Wearable vital signs monitoring for patients with asthma:
a review,'' 2020.

\bibitem{cho2021}
J.~Cho et~al., ``Wearable data quality: a systematic review,'' 2021.

\bibitem{molfino2024}
N.~A.~Molfino and G.~Turcatel, ``Machine learning approaches to predict asthma
exacerbations: a narrative review,'' 2024.

\bibitem{dehond2022}
A.~A.~H.~de~Hond et~al., ``Machine learning did not beat logistic regression
in time series prediction for severe asthma exacerbations,'' \emph{Scientific
Reports}, vol.~12, Art.~no.~20363, 2022.

\bibitem{lugogo2022}
N.~L.~Lugogo et~al., ``A predictive machine learning tool for asthma
exacerbations using an electronic inhaler with integrated sensors,''
\emph{Journal of Asthma and Allergy}, 2022.

\bibitem{tsang2023}
K.~C.~H.~Tsang et~al., ``Home monitoring with connected mobile devices for
asthma attack prediction with machine learning,'' \emph{Scientific Data},
2023.

\bibitem{chan2024}
A.~H.~Y.~Chan et~al., ``DIGIPREDICT: physiological, behavioural and
environmental predictors of asthma attacks---study protocol,'' \emph{BMJ Open
Respiratory Research}, 2024.

\bibitem{datashare}
University of Edinburgh DataShare, ``AAMOS-00 dataset repository,'' CC~BY~4.0.

\bibitem{camargo2009}
C.~A.~Camargo~Jr., G.~Rachelefsky, and M.~Schatz, ``Managing asthma
exacerbations in the emergency department,'' \emph{Proc.\ Am.\ Thorac.\
Soc.}, 2009.

\bibitem{ke2017}
G.~Ke et~al., ``LightGBM: A highly efficient gradient boosting decision
tree,'' \emph{NeurIPS}, 2017.

\bibitem{bai2018}
S.~Bai, J.~Z.~Kolter, and V.~Koltun, ``An empirical evaluation of generic
convolutional and recurrent networks for sequence modeling,'' 2018.

\bibitem{zhou2021}
H.~Zhou et~al., ``Informer: Beyond efficient transformer for long sequence
time-series forecasting,'' \emph{AAAI}, 2021.

\bibitem{lim2021}
B.~Lim et~al., ``Temporal fusion transformers for interpretable multi-horizon
time series forecasting,'' \emph{International Journal of Forecasting}, 2021.

\bibitem{ganaie2021}
M.~A.~Ganaie et~al., ``Ensemble deep learning: a review,'' 2021.

\bibitem{synapse}
Synapse, ``Asthma Mobile Health Study (AMHS),'' dataset record.

\bibitem{jozefowicz2015}
R.~Jozefowicz, W.~Zaremba, and I.~Sutskever, ``An empirical exploration of
recurrent network architectures,'' \emph{ICML}, 2015.

\end{thebibliography}

\end{document}
